Project Description -- Erosion, Transport, Anomalous Diffusion, and Collapse in Granular and Porous Media

### 2019-11-11 docs/proposals/NSF/ProjDesc/ProjDesc.tex
NM
little more stuff on transport

 \vsp{5} HERE

@@

In an eroded geometry $\Omega$, a concentration non-dimensionalized
concentration $c$ is governed by
\begin{align}
  \pd{c}{t} + \uu \cdot \nabla c + \frac{1}{\Pe} \Delta c, 
\end{align}
where $\Pe$ is the dimensionless Peclet number. Here we assume that the
concentration $c$ is affected by the eroded geometry, but the geometry
is not altered further by the concentration. In the infinite Peclet
number limit, the concentration is governed by pure advection and the
characteristics of the concentration field are identical to the
streamlines (Figure~\ref{fig:100tracers}).
\begin{wrapfigure}[]{l}{0.4\textwidth}
  \includegraphics[width=0.4\textwidth]{figs/tort_eulerian100}
  \caption{\label{fig:100tortuosity} The tortuosity of an eroding
  geometry. The tortuosity can be computed using two different formulas
  (red vs.~blue marks). The red square is the tortuosity in
  Figure~\ref{fig:100tracers}. Their is qualitative agreement between
  our numerical simulations and a power-law model (black dashed
  line)~\cite{mat-kha-koz2008}.}
\end{wrapfigure}

@@

While the tortuosity characterizes the amount of winding of streamlines,
the dispersion characterizes the spreading of the trajectories.
Dispersion can be descried using the first- and second-ensemble moments
of the trajectories
\begin{align}
  \langle \lambda \rangle (t) = \frac{1}{N_p} \sum_{j=1}^{N_p}
    \lambda_j(t), \quad \sigma_\lambda^{2}(t) = \frac{1}{N_p}
    \sum_{j=1}^{N_p} (\lambda_j(t) - \langle \lambda \rangle(t))^2
\end{align}

>>>>

 \vsp{5}
 \todo[inline]{Nick is here}

@@

In an eroded geometry $\Omega$, a concentration non-dimensionalized
concentration $c$ is governed by
\begin{align}
  \pd{c}{t} + \uu \cdot \nabla c = \frac{1}{\Pe} \Delta c, 
  \label{eqn:advectionDiffusion}
\end{align}
where $\Pe$ is the dimensionless Peclet number. Here we assume that the
concentration $c$ is affected by the eroded geometry, but the geometry
is not altered further by the concentration. In the infinite Peclet
number limit, the concentration is governed by pure advection and the
characteristics of the concentration field are identical to the
streamlines (Figure~\ref{fig:100tracers}).
\begin{wrapfigure}[]{l}{0.4\textwidth}
  \includegraphics[width=0.4\textwidth]{figs/tort_eulerian100}
  \caption{\label{fig:100tortuosity} The tortuosity of an eroding
  geometry. The tortuosity can be computed using two different formulas
  (red vs.~blue marks). The red square is the tortuosity in
  Figure~\ref{fig:100tracers}. Their is qualitative agreement between
  our numerical simulations and a power-law model (black dashed
  line)~\cite{mat-kha-koz2008}.}
\end{wrapfigure}

@@

While the tortuosity characterizes the amount of winding of streamlines,
the dispersion characterizes the spreading of the trajectories.
Dispersion can be descried using the first- and second-ensemble moments
of the trajectories
\begin{align}
  \langle \lambda \rangle (t) = \frac{1}{N_p} \sum_{j=1}^{N_p}
    \lambda_j(t), \quad \sigma_\lambda^{2}(t) = \frac{1}{N_p}
    \sum_{j=1}^{N_p} (\lambda_j(t) - \langle \lambda \rangle(t))^2
\end{align}
Then, the standard deviation $\sigma_\lambda$ characterizes the
dispersion, and long-time behavior typically results in a power-law
scaling $\sigma_\lambda \sim t^{\alpha}$. The power $\alpha$, which
quantifies the long-time spreading, is $1$ in a ballistic regime which
occurs at early times when the trajectories have not explored much of
the porous geometry. However, once the trajectories have traversed a few
grains, dispersion in porous media is often super-dispersive with
$\alpha \in (0.5,1)$~\cite{}. In Figure~\ref{fig:100dispersion}, we
investigate spreading at several different porosities of a porous
geometry that is initialized with 100 grains.

\begin{wrapfigure}[]{r}{0.5\textwidth}
  \includegraphics[width=0.5\textwidth]{figs/100b_second_moment_long_ref}
  \caption{\label{fig:100dispersion} The particle spreading,
  characterized with the standard deviation $\sigma_\lambda(t)$ at 6
  different porosities. At early times, the line of slope 1 (black
  curve) indicates a ballistic motion. Then, at later times, the
  dispersion becomes super-dispersive at a rate indicated by the slope
  of the black dashed-dotted line.}
\end{wrapfigure}

We propose to extend our preliminary results for pure advection so that
concentrations with finite Peclet numbers can be investigated.
Numerical methods to solve the advection-diffusion
equation~\eqref{eqn:advectionDiffusion} in a complex geometry poses
several challenges that we will address.  We will start by splitting the
advection and diffusion equations using a Strang splitting~\cite{}.
Numerical solutions of the advection equation will be formed with a
semi-Lagrangian method. These methods have the advantage that they ...?
The forced diffusion equation is more challenging to solve in a complex
geometry with an integral equation formulation. One particular challenge
is extending the forcing function from the complex geometry and into a
regular geometry, such as a square, so that a high-order volume integral
or Fourier method can be applied. Recently, groups have used radial
basis functions to form a sufficiently smooth extension~\cite{}
\todo[inline]{A-K, MC}. Alternatively, others have used the space-time
heat kernel~\cite{}, but this requires storing an entire time history of
the solution. We propose to develop a new method that uses the Laplace
transform.
